---
title: "PS1_ExploringData"
author: "Ji H. Kang"
date: "10/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.1 Computing 

### Question 1: Compute how many seconds are there in a year and assign it to a suitable variable. Thereafter print the result. 
```{r}
SIH <- 60 * 60 #Seconds in an hour. 60 minutes in an hour
cat("There are", prettyNum(SIH, big.mark=",", scientific=FALSE), "seconds in an hour.\n")
SID <- SIH * 24 #Seconds in a day. #24 hours in a day
cat("There are", prettyNum(SID, big.mark=",", scientific=FALSE), "seconds in a day.\n")
SIW <- SID * 7 #Seconds in a week. 7 Days in a week
cat("There are", prettyNum(SIW, big.mark=",", scientific=FALSE), "seconds in a week.\n")
SIY <- SIW * 52 #Seconds in a year. 52 weeks in a year 
cat("There are", prettyNum(SIY, big.mark=",", scientific=FALSE), "seconds in a year.\n")
```

### Question 2: How long is a typical human lifetime in seconds? Use the seconds-in-year variable you created above to compute it. 
```{r}
cat("The average human lifespan is 79 years according to google...\n")
SILT <- SIY * 79 #Seconds-in-lifetime
cat("There are approximately", prettyNum(SILT, big.mark=",", scientific=FALSE), "seconds in a human lifetime.\n")
```

### Question 3: The age of the Universe is 13.7 billion years. How old is the Universe in seconds?
```{r}
universeAge <- 13700000000 #13.7 billion 
universeAgeInSeconds <- universeAge * SILT 
cat("The Universe is", prettyNum(universeAgeInSeconds, big.mark=",", scientific=FALSE), "seconds old.\n")
```

# 1.2 Functions 

### Question 1: Write a function that takes two arguments: first name, and last name; and returns a sentence like "Hi, my name is *first name* *last name*, nice to meet you!". 
```{r}
greetPerson <- function(first_name="", last_name="") {
  return(sprintf("Hi, my name is %s %s, nice to meet you!", first_name, last_name))
}
print(greetPerson("Ji-Hoon", "Kang"))
```


# 1.3 Vectors, loops, if/else 
### Setting up the vector 
```{r}
set.seed(1) 
v <- sample(10, 20, replace=TRUE) - 5 
v
```
### Question 1: Use a for-loop to extract only positive numbers from this vector 
```{r}
positiveV <- c() #Empty vector 
for (number in v){
  if (number > 0) {
    positiveV <- c(positiveV, number)
  }
}
positiveV
```
### Not part of the assignment but can also do something like this 
```{r}
newV <- v[which(v > 0)]
newV
```

### Question 2: Perform the same task without a for-loop using logical indexing 
```{r}
logicalIndexes <- v > 0 
positiveV <- v[logicalIndexes] #Overwrite the previous vector
positiveV
```

```{r}
v1 <- 9
v2 <- c(1,2)
v3 <- c(2,3,-4)
```
### Question 3: Write a function that tests if the vectors have negative elements, and prints an appropriate message. Test the negativity of these three vectors to show the function works correctly. 
```{r}
containsNegative <- function(vec = NULL){
  if (is.null(vec)){
    print("No given vector or was NULL. Please try with a diff argument")
  } else {
    if (all(vec >= 0)){
      print("There are no negative values in this vector!")
    } else {
      print("There are negative values in this vector!")
    }
  }
}
print(paste("v1 contains:", v1))
containsNegative(v1) 
print(paste("v2 contains:", v2))
containsNegative(v2) 
print(paste("v3 contains:", v3))
containsNegative(v3) 
```

# 2.1 Exploring the NYC Flights Data 

