---
title: 'Lab 5: Probability'
author: "Ji H. Kang"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(ggplot2)
```


# Toss two coins 
Generating own data: Tossing two quarters, there were 3 different outcomes. 
0 heads & 2 tails, 1 head & 1 tail, and 2 heads & 0 tails. 
Flipping two coins 20 times, the following data was gathered: 
Number of heads per trial: 1, 2, 1, 1, 1, 0, 1, 1, 1, 1, 2, 1, 0, 1, 0, 1, 0, 0, 2, 1
Next I format the data into a readable table... 
```{r}
trial_results <- c(1, 2, 1, 1, 1, 0, 1, 1, 1, 1, 2, 1, 0, 1, 0, 1, 0, 0, 2, 1)
coin_tosses <- data.frame("Trial" = 1:20, "Number_of_Heads" = trial_results)
cat("The data generated, formatted in a table, looks like:")
knitr::kable(coin_tosses, "simple")
```

## Summary of the results 
```{r}
outcomes <- as.data.frame(table(trial_results))
names(outcomes)[1] <- "Number_of_heads"
outcomes
```
This table shows how many of the various outcomes possible occurred. 


## Computing the means of the results  
```{r}
coin_mean <- mean(trial_results)
coin_mean
```
A value of **0.9** indicates that, on average, I got less than 1 head per trial (which flips 2 coins).

## Computing the variance of the result with mean = 0.9 
Using the formula we've gone over in class... 
It is the average of the x values squared minus the value of the mean squared. 
For a refresher, this was the data: 
1, 2, 1, 1, 1, 0, 1, 1, 1, 1, 2, 1, 0, 1, 0, 1, 0, 0, 2, 1
Squaring each value we get: 
1, 4, 1, 1, 1, 0, 1, 1, 1, 1, 4, 1, 0, 1, 0, 1, 0, 0, 4, 1
Next, can quickly get the mean using R... 
```{r}
x2_mean <- mean(c(1, 4, 1, 1, 1, 0, 1, 1, 1, 1, 4, 1, 0, 1, 0, 1, 0, 0, 4, 1))
x2_mean
```
Our "x-squared-mean" has a value of 1.2 
Need to now calculate the mean squared which is simple... 
```{r}
mean_squared <- coin_mean^2
mean_squared
```
We get a value of 0.81. So doing a simple subtraction gives us... 
```{r}
coin_variance <- x2_mean - mean_squared 
cat("The variance is:", coin_variance)
```

# Plotting a histogram of the results 
```{r}
hist(trial_results, breaks=3, col="lightblue")
```

## What type of distribution is this? Do you know how it is called? 
Since the number of possible outcomes is limited to 3 different outcomes for this lab: 
* Outcomes 1: 0 heads 
* Outcomes 2: 1 head
* Outcomes 3: 2 heads
Since it can only be one of these three outcomes, it is a discrete distribution.
I'm not entirely sure what "Do you know how it is called?" is referring to but if it's asking for what kind of distribution it is, I would say it's a **binomial** distribution. I would've said it was a Bernoulli's but since there are 3 variables (0, 1, or 2 heads), it cannot be. 

## Does it look symmetric, or right/left skewed? 
Minding the gap in the histogram, it looks symmetrical. With the largest amount centered with 1 head. 