---
title: "in-class03b"
author: "Ji H. Kang"
date: "10/14/2020"
output: html_document
---
# ggplot 
```{r, message=FALSE}
library(ggplot2)
library(dplyr)
```

## Diamonds data
```{r}
data(diamonds)

```

```{r}
d1k <- diamonds %>% #1000 diamonds  
  sample_n(1000)
#Arguments are 1. Data, 2. Aesthetics mapping,
ggplot(data=d1k, aes(x=carat, y=price)) +
  geom_point() +
  #geom_line()
  
```

# Aesthetics 
**Visual properties of the plot**

Examples: 
* Horizontal position (x) 
* Vertical position (y) 
* Outline color (col)
* Fill color (fill) 
* Point size/line width (size) 
* Transparency (alpha) 
* Line type (linestyle) - maybe
* ...
```{r}
ggplot(d1k, aes(x=carat, y=price, col=depth)) +
  geom_point()
```
```{r}
ggplot(d1k, aes(x=carat, y=price, col=cut)) +
  geom_point()
```

Map one variable (price) to two aesthetics

```{r}
ggplot(d1k, aes(x=carat, y=price, col=price)) + 
  geom_point()
```

```{r}
data(Orange) 
names(Orange)
```

* Plot circumference as a function of age 
* Make point color dependent on treeid 

```{r}
ggplot(Orange, aes(age, circumference, col=Tree)) + 
  geom_point()
```

```{r}
ggplot(d1k, aes(carat, price)) + 
  geom_point(col="blue", size=3, alpha=0.1) 
```


Exercise: 
* Plot orange trees with black dots 
* Add colored lines as you did above 
```{r}
ggplot(Orange, aes(age, circumference, col=Tree, group=Tree)) + 
  geom_point(col="black") + 
  geom_line(col="lightblue")
```